---
title: 'Registration review: Registry mentions'
author: "Stylianos Serghiou"
date: '`r format(Sys.time(), "%d/%m/%Y")`'
output:
  prettydoc::html_pretty:
    # no code_folding available
    theme: hpstr      # or: architect; https://github.com/yixuan/prettydoc
    highlight: github # or: vignette
    toc: TRUE         # no toc_float available
    df_print: kable   # obviates %>% kable; does not replace styling though
  tufte::tufte_handout: default
  pdf_document:
    highlight: tango
    df_print: kable
    latex_engine: pdflatex
    keep_tex: yes
  rmdformats::readthedown:
    highlight: kate
    df_print: kable    # obviates %>% kable; does not replace styling though
    code_folding: hide # or: show; (comment out to not give option)
  tufte::tufte_html: 
    toc: TRUE
  epuRate::epurate:
    df_print: kable
    toc: yes
  html_document:
    highlight: tango
    theme: sandstone
    df_print: kable
    toc: yes
    toc_depth: 2
    toc_float: yes
    css: "path_to_custom.css"
header-includes:
- \DeclareUnicodeCharacter{3B8}{~}
- \DeclareUnicodeCharacter{3B1}{~}
- \DeclareUnicodeCharacter{3B2}{~}
- \DeclareUnicodeCharacter{223C}{~}
- \DeclareUnicodeCharacter{2264}{~}
- \DeclareUnicodeCharacter{2265}{~}
- \DeclareUnicodeCharacter{2581}{~}
- \DeclareUnicodeCharacter{2582}{~}
- \DeclareUnicodeCharacter{2583}{~}
- \DeclareUnicodeCharacter{2585}{~}
- \DeclareUnicodeCharacter{2586}{~}
- \DeclareUnicodeCharacter{2587}{~} 
- \DeclareUnicodeCharacter{FB00}{~} 
- \usepackage{graphicx}
editor_options: 
  chunk_output_type: inline
---

<style>
p {

text-align: justify;
text-justify: interword;
padding: 0 0 0.5em 0

}
</style>

```{r setup, include=FALSE}
# Load packages
library(knitr)
library(rmdformats)
library(kableExtra)
library(ggplot2)
library(magrittr)



######### knitr

# Define chunk options
opts_chunk$set(
  echo = T,
  cache = F,  # if TRUE, no need to rerun chunks
  # cache.lazy = TRUE,  # use with big objects (>1 GB)
  cache.comments = F,  # do not rebuild if comments change
  tidy = F,  # can play with this
  warning = F, 
  message = F,
  comment = NA,
  fig.align = "center",
  fig.width = 7,
  fig.path = "Figs/",  # export all figures to dir Figs
  linewidth = 91,
  width = 75
)


# Initiatialize hook
hook_output = knit_hooks$get("output")


# Hook to wrap output text when it exceeds 'n' using linewidth
knit_hooks$set(output = function(x, options) {
  
  if (!is.null(n <- options$linewidth)) {
    x = knitr:::split_lines(x)
    
    # wrap lines wider than 'n' 
    if (any(nchar(x) > n)) 
      x <- strwrap(x, width = n)
      x <- paste(x, collapse = "\n")
  }
  hook_output(x, options)
})


# Times a chunk and prints the time it took to run it under the chunk
# To time a chunk, include in the chunk options: {r my_chunk, timeit=TRUE}
knitr::knit_hooks$set(timeit = local({
  now = NULL
  function(before, options) {
    if (before) {
      now <<- Sys.time()
    } else {
      res = difftime(Sys.time(), now)
      now <<- NULL
      # use options$label if you want the chunk label as well
      paste('Time for this code chunk:', as.character(res))
    }
  }})
)


# For more knitr options visit: https://yihui.name/knitr/options/
# and his github page: https://github.com/yihui/knitr-examples



######### kableExtra

options(knitr.kable.NA = ''  # replace NAs in tables with blank
        , digits = 3)          # round digits (doesn't work without this here!)

## Example use
# data.frame(x = c(1,2,3), y = c(4,5,6), z = c(7,8,9)) %>% 
#   kable(booktabs = T) %>% kable_styling()

# Function to simplify table styling
sable <- function(tab, escape = T, full_width = F, drop = F, font_size = 12) {
  if (drop) {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      collapse_rows(valign = "top") %>% 
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width, 
                    font_size = font_size)
  } else {
    tab %>%
      kable(escape = escape, booktabs = T) %>%
      kable_styling("striped", 
                    position = "center", 
                    full_width = full_width,
                    font_size = font_size)
  }
}

## Guidelines
# No longer need to define options(knitr.table.format = "html"). It is now automatically done as soon as you load kableExtra
# No need to run kable() every time - done automatically as soon as you load kableExtra
# Loading kableExtra nullifies any styling applied by df_table: kable in the preamble - if you are content with standard formatting, DO NOT load kableExtra



#########  ggplot2

# Set up preferred theme in ggplot2
my_theme <- 
  # this and theme_minimal() are my favorite
  theme_light() +  
  theme(
    axis.ticks = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.title.x = element_text(margin = margin(t = 15)),
    axis.title.y  = element_text(margin = margin(r = 7)),
    legend.key = element_rect(colour = NA, fill = NA),  # Avoid borders
    panel.border = element_blank(), 
    text = element_text(color = "grey20"),
    title = element_text(face = "bold")
  )

# Make the above theme the default theme
original_theme <- theme_set(my_theme)

# Use ggsave to save plots after plotting - this reduces size dramatically



######### Live preview

# Preview the HTML without having to re-knit
# https://www.r-bloggers.com/create-r-markdown-reports-and-presentations-even-better-with-these-3-practical-tips/amp/

# xaringan::infinite_moon_reader()




######### Tabbed sections

# You can organize content using tabs by applying the .tabset class attribute to headers within a document. This will cause all sub-headers of the header with the .tabset attribute to appear within tabs rather than as standalone sections. For example:

## Quarterly Results {.tabset}

### By Product



######### Update package

# To update the package use:
# Replace ~/serghiouTemplates/inst/rmarkdown/templates/report/skeleton.rmd
# library(devtools); setwd("/Users/Stelios/"); install("serghiouTemplates")
```


# Setup {.tabset}

<!-- Chunk called setup_2 to avoid running it automatically -->

```{r setup_2}
# Load packages
library(magrittr)
library(readxl)
library(tidyverse)
library(vroom)

# Import data
pmcoa <- vroom("../../data/tidy_data/pmcoa.csv")
```



***


# Function

Define functions to identify all [registries monitored by the WHO](https://www.who.int/ictrp/search/data_providers/en/), as well as PROSPERO. Notice that many of the regular expressions could be simplified, but were not to allow for flexibility in possible future uses of this code.

```{r}
# Dictionary of registry regex strings
registry_dic <- function() {
  
  list(
    
    ANZCTR = c(
      "ANZCTR",
      "Zealand Clinical Trial",
      "ACTR.\\s*[0-9]{14}"
    ),
    
    ChiCTR = c(
      "ChiCTR",
      "(?i)Chinese Clinical Trial(|s) Register(?-i)",
      "ChiCTR\\s*[0-9]{10}"
    ),
    
    ClinicalTrials = c(
      "[Cc]linical[Tt]rials",
      "NCT\\s*[0-9]{8}"
    ),
    
    CRIS = c(
      "\\bCR[iI]S\\b",
      "KCT\\s*[0-9]{7}"
    ),
    
    CTRI = c(
      "Clinical Trial(|s) Registry( |- )India",
      "\\bCTRI\\b"
    ),
    
    DRKS = c(
      "\\bDRKS\\b",
      "German Clinical Trial(|s) [Rr]egister",
      "DRKS.{0,5}[0-9]{8}"
    ),
    
    EudraCT = c(
      "\\bE(?i)udraCT(?-i)\\b"
    ),
    
    EUCTR = c(
      "\\bEU(-| )CTR\\b",
      "EU Clinical Trial(|s) [Rr]egister"
    ),
    
    EUPAS = c(
      "\\bEU PAS\\b",
      "EUPAS\\s*[0-9]{5}"
    ),
    
    IRCT = c(
      "\\bIRCT\\b",
      "Iranian Registry of Clinical Trial(|s)",
      "IRCT\\s*[0-9]{12}"
    ),
    
    ISRCTN = c(
      "\\bISRCTN\\b",
      "\\bISCRTN\\b",
      "\\bICTRP\\b",
      "International Clinical Trial(|s) Registry",
      "ISRCTN\\s*[0-9]{8}",
      "ISCRTN\\s*[0-9]{8}",
      "ICTRP\\s*[0-9]{5}"
    ),
    
    JAPIC = c(
      "\\bJAPIC\\b",
      "CTI.{0,3}[0-9]{6}"
    ),
    
    JPRN = c(
      "\\bJPRN\\b",
      "\\bUMIN\\b",
      "Japan Primary Registries Network",
      "UMIN.{0,5}[0-9]{9}"
    ),
    
    NTR = c(
      "Netherlands National Trial(|s) [Rr]egister",
      "NTR.{0,5}[0-9]{4}",
      "Nederlands Trial(|s) [Rr]egister [0-9]{4}"
    ),
    
    OSF = c(
      "\\bOSF\\b",
      "Open Science Framework"
    ),
    
    PACTR = c(
      "\\bPACTR\\b", 
      "Pan African Clinical Trial",
      "PACTR\\s*[0-9]{15}"
    ),
    
    PROSPERO = c(
      "\\bP(?i)ROSPERO(?-i)\\b",
      "CRD\\s*[0-9]{11}"
    ),
    
    ReBec = c(
      "\\bReBec\\b",
      "Brazilian Clinical Trial(|s)",
      "RBR-[0-9a-zA-Z]{6}"
    ),
    
    REPEC = c(
      "\\bREPEC\\b",
      "Peruvian Clinical Trial(|s) [Rr]egistry"
    ),
    
    RPCEC = c(
      "\\bRPCEC\\b",
      "Cuban Public Registry of Clinical Trials"
    ),
    
    SLCTR = c(
      "\\bSLCTR\\b",
      "Sri Lanka Clinical Trial(|s) [Rr]egistry"
    ),
    
    TCTR = c(
      "\\bTCTR\\b",
      "Thai Clinical Trial(|s) Register"
    )
  )
  
}
```


```{r}
#' Identify registries within text
#' @param x A string.
#' @return A string of the registries identified.
extract_registry <- function(x) {
  
  # Return NA if NA
  if (is.na(x)) {
    
    return(NA_character_)
    
  }
  
  # Pre-process text
  x %<>% 
    str_replace_all("[[:punct:]]", " ") %>% 
    str_squish()
  
  # Create regex
  registries <- lapply(registry_dic(), paste, collapse = "|")
    
  # Detect registries
  registries %>% 
    purrr::map(str_detect, string = x) %>%
    purrr::keep(~ .x) %>% 
    names() %>% 
    paste(collapse = "; ")
}
```


***


# Run

```{r, eval=FALSE}
registry_mentions <- 
  pmcoa %>%
  filter(is_register_pred) %>% 
  rowwise() %>% 
  mutate(registries = extract_registry(register_text)) %>%
  select(pmcid_pmc, registries) %>% 
  ungroup()
```


***


# Performance

Test performance in 50 random samples.

```{r, eval=FALSE}
# Set seed
set.seed(2115)

# Test
registry_mentions %>% 
  filter(nchar(registries) == 0) %>% 
  select(register_text) %>% 
  sample_n(50) %>%
  write_csv("../../data/interim_data/pmcoa_registry-test.csv")
```


Describe performance.

```{r}
# Import test
registry_test <- read_xlsx("../../data/interim_data/pmcoa_registry-test.xlsx")

# Summarize test
registry_test %>% 
  count(is_true_negative, reason, sort = T) %>%
  group_by(is_true_negative) %>% 
  mutate(N = sum(n)) %>% 
  mutate(p = scales::percent(n / N, accuracy = 0.1)) %>%
  arrange(desc(is_true_negative)) %>% 
  sable()
```

Out of 50 random records predicted to not share any registry, 37 truly did not (TN), but 13 did (FN). Of 13 FNs, 8 referred to a register that the code was not designed to capture (e.g. CAMARADES, Cochrane), 3 referred to an unknown or possibly false registry and 2 referred to registers that the code was meant to capture (ANZCTR, EudraCT). None of the registers that the code was not designed to capture appeared more than once. Projecting to 7318 articles predicted to not contain a recognized register, we expect that this code is missing 7318 / 50 * 2 = 293 of the registers it was designed to capture, which does not meaningfully alter our results. 


***


# Save

```{r save, eval=FALSE}
# Save
write_csv(registry_mentions, "../../data/tidy_data/registry_mentions.csv")
```


***


# Documentation {.tabset}

## Session Info

```{r session_info, echo=FALSE}
print(sessionInfo(), locale = F)
```


## References

```{r refs, echo=FALSE}
(.packages()) %>% sort %>% lapply(citation) %>% lapply(c) %>% unique
```
